---
title: "Production Performance"
author: "Miguel Duarte B."
date: "5/9/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Hydraulic fracturing is a process applied to unconventional reservoirs to create
artificial fractures connecting the pores where hydrocarbons are stored. The
estimation of formation properties such as permeability and effective half
fracture length is an important step in analyzing production performance.

Luo et al., 2016, developed a method to calculate permeability and effective
half length using the square root plot of the production.

## Methodology

1. Select the well for analysis from the drop-down menu
2. Select the end of the linear flow using the slider. The idea is to fit the
blue line (linear model) as close as possible to straight line formed by the
black dots.  This is equivalent to estimate visually the end of linear flow (red
line) and pass this value to the model
3. If necessary adjust the max value of Y using the slider
4. The estimated values will be shown at the bottom of the plot

Check the app: [Production Performance Shiny App](https://migueldb.shinyapps.io/production-performance/)

## Square Root of Time Plot

```{r prepro, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
## global.R
## This file is run once before the app starts.  All the objects created here
## are available for ui.R and server.R
# 

# Make sure the environment is clean
rm(list = ls())

# Load libraries
library(tidyverse)

# Conversion factors
feetToMeters <- 1200/3937
psiToMPa <- 0.00689475729
stbTom3 <- 0.1589872943

# # Make sure this is the right folder
# print(getwd())

# Find and read the source files
folderPath <- file.path(getwd(), "production-performance/Raw/SPE")
productionFile <- file.path(folderPath, "production_data.txt")
wellDataFile <- file.path(folderPath, "well_data.txt")

# Read production
production <- read_csv(
    productionFile,
    col_types = cols(
        Lease = col_character(),
        `Pressure Source` = col_character(),
        .default = col_double()
    ))
# Remove double space from column names
names(production) <- gsub("  \\(", " \\(", names(production))

# Read well data
wellData <- read_csv(
    wellDataFile,
    col_names = FALSE,
    col_types = cols(.default = col_character())
    )

# Need to transform "transposed" data from csv file
wellData <- wellData %>% t() %>% as_tibble(.name_repair = "unique")

names(wellData) <- wellData %>% slice(1) %>% as_vector() %>% make.unique()
wellData <- wellData %>% slice(-1)

# Remove columns with only NA values
not_all_na <- function(x) any(!is.na(x))
not_any_na <- function(x) all(!is.na(x))

production <- production %>% select(where(not_all_na))
wellData <- wellData %>%
    # Remove variables with only NA values
    select(where(not_all_na)) %>%
    # Parse column types
    mutate(across(everything(), parse_guess)) %>%
    # parse character columns with percentages
    mutate_if(~any(str_detect(.,"\\d+\\.?\\d*%")), ~parse_number(.)/100) %>%
    # Mutate If has been superseded with mutate across where combination but
    # straight conversion doesn't work.  Check:
    # https://dplyr.tidyverse.org/articles/colwise.html#how-do-you-convert-existing-code-
    #mutate(across(where(~any(str_detect(.x ,"\\d+\\.?\\d*%"))), ~parse_number(.x)/100))
    mutate(`# of Total Proppant (Lbs)` = parse_number(`# of Total Proppant (Lbs)`))

# Summarize production data by lease
MMscf_to_MMboe <- 1/5800.6408

productionSummary <- production %>%
    group_by(Lease) %>%
    summarise(
        `Cummulative Oil Volume (stb)` = sum(`Oil Volume (stb)`, na.rm = TRUE),
        `Cummulative Gas Volume (MMscf)` = sum(`Gas Volume (MMscf)`, na.rm = TRUE),
        `Cummulative Water Volume (stb)` = sum(`Water Volume (stb)`, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(
        `Cummulative Hydrocarbons Volume (MMboe)` =
            `Cummulative Oil Volume (stb)` +
            `Cummulative Gas Volume (MMscf)` *MMscf_to_MMboe
    )

# # Summarize production data by lease - 1st 100 days
# productionSummary100 <- production %>%
#     group_by(Lease) %>%
#     slice(1:100) %>%
#     summarise(
#         `Cummulative Oil Volume (stb) - 100 days` = sum(`Oil Volume (stb)`, na.rm = TRUE),
#         `Cummulative Gas Volume (MMscf) - 100 days` = sum(`Gas Volume (MMscf)`, na.rm = TRUE),
#         `Cummulative Water Volume (stb) - 100 days` = sum(`Water Volume (stb)`, na.rm = TRUE)
#     ) %>%
#     ungroup() %>%
#     mutate(
#         `Cummulative Hydrocarbons Volume (MMboe) - 100 days` =
#             `Cummulative Oil Volume (stb) - 100 days` +
#             `Cummulative Gas Volume (MMscf) - 100 days` * MMscf_to_MMboe
#     )
# 
# # Summarize production data by lease - 1st 200 days
# productionSummary200 <- production %>%
#     group_by(Lease) %>%
#     slice(1:200) %>%
#     summarise(
#         `Cummulative Oil Volume (stb) - 200 days` = sum(`Oil Volume (stb)`, na.rm = TRUE),
#         `Cummulative Gas Volume (MMscf) - 200 days` = sum(`Gas Volume (MMscf)`, na.rm = TRUE),
#         `Cummulative Water Volume (stb) - 200 days` = sum(`Water Volume (stb)`, na.rm = TRUE)
#     ) %>%
#     ungroup() %>%
#     mutate(
#         `Cummulative Hydrocarbons Volume (MMboe) - 200 days` =
#             `Cummulative Oil Volume (stb) - 200 days` +
#             `Cummulative Gas Volume (MMscf) - 200 days` * MMscf_to_MMboe
#     )


# From well data filter out unnecessary data
wellDataSummary <- wellData %>%
    select(
        Lease,
        `Formation/Reservoir`,
        `Initial Pressure Estimate (psi)`,
        `Reservoir Temperature (deg F)`,
        `Net Pay (ft)`,
        `Wellbore Diameter (ft)`,
        Porosity,
        `Oil Gravity (API)`,
        `# Stages`,
        `# Clusters`,
        `# Clusters per Stage`,
        `# of Total Proppant (Lbs)`,
        `Total Fluid (Bbls)`,
        `Lateral Length (ft)`,
        `Sandface Temp (deg F)`
    ) %>%
    # Calculate the necessary variables
    mutate(
        `Spacing (ft)` =  `Lateral Length (ft)` / `# Stages`,
        `Spacing (m)` = `Spacing (ft)` * feetToMeters,
        `Lateral Length (m)` = `Lateral Length (ft)` * feetToMeters,
        `Net Pay (m)` = `Net Pay (ft)` * feetToMeters,
        # Viscosity calculations based on Bennison correlation
        # https://petrowiki.spe.org/Oil_viscosity
        `Viscosity (mPa s)` = 10^(-0.8021*`Oil Gravity (API)` + 23.8765)*`Reservoir Temperature (deg F)`^(0.31458*`Oil Gravity (API)` - 9.21592),
        `Total Compressibility (MPa -1)` = 0.004,
        `Payzone Thickness (m)` = `Net Pay (ft)` * feetToMeters
        )

# Combine production summary and well data
# extendedWellData <- inner_join(productionSummary100, productionSummary200, by = 'Lease') %>%
#     inner_join(productionSummary, by = 'Lease') %>%
#     inner_join(wellDataSummary, by = 'Lease') %>%
extendedWellData <- inner_join(productionSummary, wellDataSummary, by = 'Lease') %>%
    # Create GOR variable and oil or gas well designation based on 100,000 scf/bbl limit
    mutate(`GOR scf/bbl` = `Cummulative Gas Volume (MMscf)` * 1e6 / `Cummulative Oil Volume (stb)`) %>%
    mutate(Designation = if_else(`GOR scf/bbl` >= 1e5, "Gas Well", "Oil Well"))

# Create list of oil wells based on GOR criteria
oilWells <- extendedWellData %>% filter(Designation == "Oil Well") %>% pull(Lease) %>% unique()

# Calculate necessary variables on production df and remove unnecessary gas
# wells

IRP <- function(well, df = wellDataSummary) df %>% filter(Lease == well) %>% pull(`Initial Pressure Estimate (psi)`)
production <- production %>%
    filter(Lease %in% oilWells) %>%
    mutate(
        `Calculated Sandface Pressure (MPA(a))` = `Calculated Sandface Pressure (psi(a))` * psiToMPa
    ) %>%
    rowwise() %>%
    mutate(`Initial Reservoir Pressure (psi(a))` = IRP(Lease)) %>%
    ungroup() %>%
    mutate(
        `Initial Reservoir Pressure (MPa)` = `Initial Reservoir Pressure (psi(a))` * psiToMPa,
        `Active Pressure (MPa)` = `Active Pressure (psi(a))` * psiToMPa,
        # `Drawdown Pressure (psi(a))` = `Calculated Sandface Pressure (psi(a))` - `Active Pressure (psi(a))`,
        # `Drawdown Pressure (MPa)` = `Calculated Sandface Pressure (MPA(a))` - `Active Pressure (MPa)`,
        `Drawdown Pressure (psi(a))` = `Initial Reservoir Pressure (psi(a))` - `Active Pressure (psi(a))`,
        `Drawdown Pressure (MPa)` = `Initial Reservoir Pressure (MPa)` - `Active Pressure (MPa)`,
        `Oil Volume (m3)` = `Oil Volume (stb)` * stbTom3,
        `Normalized Pressure (psi(a)/stb/day)` = if_else(`Oil Volume (stb)` == 0, 0, `Drawdown Pressure (psi(a))` / `Oil Volume (stb)`),
        `Normalized Pressure (MPa/m3/day)` = if_else(`Oil Volume (m3)` == 0, 0, `Drawdown Pressure (MPa)` / `Oil Volume (m3)`)
        )





## Define useful functions available for ui.R and server.R ####

# Root square of time plot and lm coefficients
plot.sqrt.time <- function(well, divider, df = production, maxY = 5){
    ## Load libraries
    library(tidyverse)
    library(ggplot2)
    # Filter dataframe by well and apply divider as upper limit
    df <- df %>% filter(Lease == well)
    
    df_r <- df %>% filter(`Time (Days)` <= divider^2)
    
    model <- lm(`Normalized Pressure (MPa/m3/day)` ~ sqrt(`Time (Days)`), df_r)
    intercept <- coef(model)[1]
    slope <- coef(model)[2]
    
    g <- df %>%
        ggplot(aes(sqrt(`Time (Days)`), `Normalized Pressure (MPa/m3/day)`, )) +
        geom_point() +
        geom_abline(intercept = intercept, slope = slope, colour = "blue") +
        geom_vline(xintercept = divider, colour = "red") +
        ylim(0, maxY)
    
    return(list(g, intercept, slope))
}

# Calculate permeability
permeability <- function(d, phi, mu, c_t = 0.0004, T_elf){
    79.014*d^2*phi*mu*c_t/4/T_elf
}

# Calculate permeability
frac.half.length <- function(B = 1, T_elf, m, l_e, h, phi, c_t = 0.0004){
    4.972*B*sqrt(T_elf)/m/l_e/h/phi/c_t
}
```

```{r sqrtplot, message=FALSE, warning=FALSE, dependson='prepro', fig.dim=c(5,2.5), fig.align='center'}
plot.sqrt.time("EAGLE", 20)[[1]]
```

<font size="2"> The permeability can be calculated using the limit of linear flow (red line) using the app.</font>
```{r permeability, dependson='prepro', echo=TRUE, include=TRUE}

df <- wellDataSummary %>% filter(Lease == "EAGLE")
        d <- df %>% pull(`Spacing (m)`); phi <- df %>% pull(Porosity)
        mu <- df %>% pull(`Viscosity (mPa s)`); T_elf <- 20^2
        
        permeability(d = d, phi = phi, mu = mu, T_elf = T_elf)

```

## References

- Dataset: SPE Data Repository: Data Set: 1, Well Number: all wells, From URL: [https://www.spe.org/en/industry/data-repository/](https://www.spe.org/en/industry/data-repository/)
- Luo, H., Li, H., Zhang, J. et al. Production performance analysis of fractured horizontal well in tight oil reservoir. J Petrol Explor Prod Technol 8, 229–247 (2018). [https://doi.org/10.1007/s13202-017-0339-x](https://doi.org/10.1007/s13202-017-0339-x)
- [https://github.com/migueldb/production-performance](https://github.com/migueldb/production-performance)
- [Production Performance Shiny App](https://migueldb.shinyapps.io/production-performance/)
- Author: Miguel Duarte B. [LinkedIn](https://www.linkedin.com/in/miguel-duarte-oandg/)
